---
title: "SampleBismarkMakeData"
author: "Caihong Xia"
output: html_document
---

```{r echo = FALSE, warning=FALSE}
library(dplyr)
library(data.table)

read_met <- function(x) {
  dat <- fread(x, header=F)
  colnames(dat) <- c("chr", "start", "end", "beta", "num_met", "num_unmet")
  dat$coverage <- dat$num_met + dat$num_unmet
  return(dat)
}

read_pos <- function(x){
  dat <- fread(x, header = F)
  colnames(dat) <- c('Bismark', 'methylation', 'extractor', 'version', 'v0.22.3')
  return(dat)
}

makedata <- function(metdata, cpgob, cpgot){
  
# C position on OB strand
cposob <- cpgob %>% distinct(extractor, version) %>% arrange(extractor, version) %>% rename("extractor"="chr", "version"="end") %>% dplyr:: select(chr, end)

# C position on OT strand
cposot <- cpgot %>% distinct(extractor, version) %>% arrange(extractor, version) %>% rename("extractor"="chr", "version"="start") %>% dplyr:: select(chr, start)

# Get metdata on covered positions
datob <- inner_join(metdata, cposob, by = c('chr','end')) %>% dplyr::select(chr, end, num_met, num_unmet)
datot <- inner_join(metdata, cposot, by = c('chr', 'start')) %>% mutate(end=start+1) %>% dplyr::select(chr, start, end, num_met, num_unmet)

# Convert C-metrics data to CpG-metrics data
datall <- full_join(datob, datot, by = c('chr', 'end')) %>% 
            mutate(start = ifelse(is.na(start), end-1, start),
                    start = as.integer(start),
                     end = as.integer(end),
                    num_met.x = ifelse(is.na(num_met.x), 0, num_met.x),
                     num_met.y = ifelse(is.na(num_met.y), 0, num_met.y),
                       num_unmet.x = ifelse(is.na(num_unmet.x), 0, num_unmet.x),
                         num_unmet.y = ifelse(is.na(num_unmet.y), 0, num_unmet.y),
                           num_met=num_met.x+num_met.y, 
                             num_unmet=num_unmet.x+num_unmet.y,
                               coverage=num_met+num_unmet,
                               beta=num_met/coverage) %>% 
              # filter(coverage>4) %>% 
                dplyr:: select(chr, start, end, beta, coverage, num_met, num_unmet) %>% 
                   arrange(chr, start, end)

return(datall)
}

```


```{r mkmetdata}

 metdata <- read_met("AS1L1.multiple.deduplicated.bismark.cov.gz")
 cpgob <- read_pos("CpG_OB_AS1L1.multiple.deduplicated.txt.gz")
 cpgot <- read_pos("CpG_OT_AS1L1.multiple.deduplicated.txt.gz")
 metdata_s1 <- makedata(metdata, cpgob, cpgot)
 #summary(metdata_s1)
 #write.table(metdata_s1, "wg_s1n.bed", row.names = F, col.names = F, quote = F, sep = "\t")

 ···
 
 # Make pwd bed file after making metdata for all 6 samples;
 
 ```{r makedata pwd ucsc}
gendf <- function(data){
#data$pwd <- (data$num_met.x*data$num_unmet.y+data$num_met.y*data$num_unmet.x)/data$coverage.x/data$coverage.y
  data$pwd <- abs(data$beta.x - data$beta.y)
  df <- data %>% filter(coverage.x>4 & coverage.y>4) %>%  dplyr:: select(chr, start, end, pwd, beta.x, beta.y)
return(df)
}

# S1-S4: N1-N2
joins <- inner_join(metdata_s1, metdata_s4, by = c('chr', 'start', 'end'))
df14pwd <- gendf(joins)
summary(df14pwd)
write.table(df14pwd, "wgdf14pwd.bed", row.names = F, col.names = F, quote = F, sep = "\t")

# S2-S3: TA1-TB1
joins <- inner_join(metdata_s2, metdata_s3, by = c('chr', 'start', 'end')) 
df23pwd <- gendf(joins)
summary(df23pwd)
write.table(df23pwd, "wgdf23pwd.bed", row.names = F, col.names = F, quote = F, sep = "\t")

# S5-S6: TA2-TB2
joins <- inner_join(metdata_s5, metdata_s6, by = c('chr', 'start', 'end'))
df56pwd <- gendf(joins)
summary(df56pwd)
write.table(df56pwd, "wgdf56pwd.bed", row.names = F, col.names = F, quote = F, sep = "\t")

```
